Interpretable Neural Networks using EAGGA

### 2025-03-14 paper/main.tex
introduction bullets

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
%\blindtext
- propose a network architecture specifically suited for the EAGGA algorithm
- achieve comparable performance as training XGBoost with EAGGA
\end{abstract}

@@

\section{Introduction}
Tabular Data
- most common type of data
- still difficult for neural networks  % TODO: reference, or better only in introduction
- recent research suggests that heavy regularisation is necessary  % TODO: reference, or better only in introduction

regularisation
- LM: ridge (L2) + lasso (L1)
  * ridge: overcome multi-collinearity by pulling coeffs close to 0
  * lasso: variable selection by setting coeffs = 0
- NN:
  * L2 regularisation implicitly included in SGD when using weight decay (pull params towards 0) -> does this combat multi-collinearity as well or serve another purpose?
  * L1
    + key claim of DL: feature selection obsolete, network finds important ones itself
    + still, feature usually always has some impact (back this up), no explicit feature selection
    + explicit feature selection simply by not inluding a feature
  * another classical NN regularisation technique: dropout against co-adaptation
  * early stopping against overfitting when using iterative method
-> all of those require specifying Hyperparameters in advance

@@

\subsection{Neural Architecture Search (NAS)}
- approaches specifically for NN HPO due to flexible structure of NNs and thus very large search space (each layer can have different \# nodes, different activations, for CNN different pooling, etc. operations)
- we don't employ NAS for our extension as research focusses on the NLP and image domains, % TODO: find some more references
  whose datasets exhibit strong correlation amongst features (i.e. tokens or pixels, respectively), an effect that is much weaker for tabular data \citep[p. 1]{Borisov_2024}
- at least want to give a little overview over most notable approaches
  * \textbf{cell search space}
    - modularise NN architecture into cells, NN as chain-structure of those
    - cell = basic building block, fixed component of an NN
      * e.g. linear layer with specific \# of neurons in feedforward
      * or convolutional / pooling / etc layer for CNN
    - then optimise sequential placement of the cells as HPO problem
      * e.g. via random search or BO
      * \citet[p. 3]{zoph2017neuralarchitecturesearchreinforcement} + \citet[pp. 2-4]{Zoph_2018_CVPR} use RNN to recursively optimise HPs of a cell given already determined
        previous cells (+ HPs of current cell), which is trained via RL with the different HPs the RNN predicts being the actions and performance on held-out data being reward
    - \citep[chap. 3.2]{elsken_neural_2019}
  * \textbf{one-shot model}
    - trains one overall ``fabric'' comprising all architectures of search space
    - visualisation as DAG, nodes are nodes, edges in-between are operations on a node
    - each path through DAG (from input to output node) represents one architecture of search space -> train entire DAG, then pick optimal path (architecture)
    - pro: very efficient training, individual architectures share operations along edges they share, training one-shot model trains all subsumed models (more expensive
      than training a single model but less expensive than trying out all configurations included in the fabric)
    - \citep[pp. 1-2, p.8]{saxena2017convolutionalneuralfabrics}
    % don't necessarily talk about co-adaptation problem

>>>>

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
%\blindtext
- tabular data still difficult for NNs, where it's still outperformed by other ML model classes
- research suggests NN performance benefits from heavy regularisation
- using EAGGA, we regularise an NN and achieve both comparable performance as XGB on EAGGA as well as interpretability
- for this, we propose a network architecture specifically suited for the EAGGA algorithm
\end{abstract}

@@

\section{Introduction}
Tabular Data
- most common type of data
- still difficult for neural networks
- \citep[p. 7499]{Borisov_2024}
- recent research suggests that strong regularisation is beneficial to NN performance on tab data \citep[8]{NEURIPS2021_c902b497}

- we know regularisation from linear models can come with improvements in interpretabiltiy
  * e.g. LM-LASSO (L1) regularisation as feature selection -> reduces \# features used in model by setting some coeffs to 0 \citep[p. 267]{lasso}
- other forms of regularisation improve performance
  * e.g. LM-Ridge (L2) on multi-collinear data \citep[p. 55]{ridge}  % TODO: mention in "extension" section that L2 is implemented via weight decay
  * e.g. NN dropout, reduces co-adaptation \citep[p. 1]{hinton2012improvingneuralnetworkspreventing}
  * e.g. NN early stopping, reduces overfitting on training data \citep[p. 778f]{early_stopping}

- we want to explore if we can use NN regularisation to tackle both interpretability and improved performance (i.e. ``comparable'' to XGBoost) on tab data
- using EAGGA framework, which proved it can improve performance while keeping (already high) performance of XGBoost on tab data -> see if interpretabiltiy
  improvements translate to NN + performance can also be on par

@@

\subsection{Neural Architecture Search (NAS)}
- approaches specifically for NN HPO due to flexible structure of NNs and thus very large search space (each layer can have different \# nodes, different activations, for CNN different pooling, etc. operations)
- we don't employ NAS for our extension as research focusses on the NLP and image domains, % TODO: find some more references
  whose datasets exhibit strong correlation amongst features (i.e. tokens or pixels, respectively), an effect that is much weaker for tabular data \citep[p. 7499]{Borisov_2024}
- at least want to give a little overview over most notable approaches
  * \textbf{cell search space}
    - modularise NN architecture into cells, NN as chain-structure of those
    - cell = basic building block, fixed component of an NN
      * e.g. linear layer with specific \# of neurons in feedforward
      * or convolutional / pooling / etc layer for CNN
    - then optimise sequential placement of the cells as HPO problem
      * e.g. via random search or BO
      * \citet[p. 3]{zoph2017neuralarchitecturesearchreinforcement} + \citet[pp. 2-4]{Zoph_2018_CVPR} use RNN to recursively optimise HPs of a cell given already determined
        previous cells (+ HPs of current cell), which is trained via RL with the different HPs the RNN predicts being the actions and performance on held-out data being reward
    - \citep[chap. 3.2]{elsken_neural_2019}
  * \textbf{one-shot model}
    - trains one overall ``fabric'' comprising all architectures of search space
    - visualisation as DAG, nodes are nodes, edges in-between are operations on a node
    - each path through DAG (from input to output node) represents one architecture of search space -> train entire DAG, then pick optimal path (architecture)
    - pro: very efficient training, individual architectures share operations along edges they share, training one-shot model trains all subsumed models (more expensive
      than training a single model but less expensive than trying out all configurations included in the fabric)
    - \citep[pp. 1-2, p.8]{saxena2017convolutionalneuralfabrics}
    % don't necessarily talk about co-adaptation problem
